%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 2 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Many-Body Quantum Theory\label{chapter:2}}
    Ever since the dawn of Quantum-Mechanics, with Boltzmanns model for the
    $I_2$ molecule in 1877 to the photoelectric by Einstein in 1905, Bohrs
    model for the hydrogen atom in 1913 and - of course - the discovery of the
    Schrödinger equation, physicists have been qurious as to if the theory
    could be applied to larger systems. The first milestone however can be
    traced back to the 1927 paper by Walter Heitler and Fritz London which
    studied the stability of the hydrogen molecule and of course the well-known
    Pauli-principle. Still it wasn't before later when increased computational
    power came into play that many-body quantum mechanics and many-body quantum
    chemistry sky-rocketed and started to become as natural a part of the
    science laboratory as physical experiments since the former is far less
    expensive\footnote{Economically}. The challenge thus became to tackle the
    ever-increasing computational cost of actually solving Schrödingers
    equation for a many-body system.

    Many methods have been devised and the basis in quantum chemistry is the
    \txtit{Hartree-Fock method} while it being a good first approximation, it
    is known to be poor at modeling correlations and is only useful for
    building a basis set to be used in other methods like
    \txtit{coupled-cluster} or \txtit{configuration interaction}. The quantum
    mechanical properties as a whole is also approximated with the
    \txtit{Born-Oppenheimer approximation} and individual electrons are not
    assigned to a single atomic core, but rather spread out across the
    molecule.

    We will in this chapter take forth the theory regarding the basics of
    identical particles and \txtit{many-body quantum mechanics}. The reader is
    referred to \cite{GriffQuan} for an introductory text on quantum
    mechanics(for single particles) and also the so-called
    \txtit{Dirac-notation} used throughout the entire chapter. We will address
    methods regarding computational quantum mechanics and further deepen into
    Hartree-Fock methods and Variational Monte Carlo method. 

\section{The Hamiltonian and the Born-Oppenheimer Approximation\label{sec:3.1}}
    The task at hand is to solve the many-body system described by
    \txtit{Schrödinger's} equation
        \begin{equation}
            \mathcal{H}\ket{\Psi_i} = E_i\ket{\Psi_i}
            \label{eq:SE}
        \end{equation}
    for some state $\ket{\Psi_i}$ with energy $E_i$. Usually the desired state
    is the ground-state energy $E_0$ of the system meaning we are primarily
    interested in the \txtit{ground-state} $\ket{\Psi_0}$. \\ With the goal
    determined we can define the system to consist of $N$ identical
    particles\footnote{These are in both atomic physics and in the quantum dot
    case always fermions or bosons.} with positions
    $\{\blds{r}_i\}^{N-1}_{i=0}$ and $A$ nuclei with positions
    $\{\blds{R}_k\}^{A-1}_{k=0}$. The Hamiltonian $H$ is then
        \begin{equation}
            \mathcal{H} = - \frac{1}{2} \sum\limits_i \nabla^2_i + \sum_{i<j}
            f\left(\blds{r}_j, \blds{r}_j\right) - \frac{1}{2} \sum_k
            \frac{\nabla^2_k}{M_k} + \sum_{k<l}
            g\left(\blds{R}_k,\blds{R}_l\right) +
            V\left(\blds{R},\blds{r}\right)
        \end{equation}
    The first and second terms represent the kinetic- and inter-particle
    interaction terms\footnote{This is usually the well-known Coulomb
    interaction.} for the $N$ identical particles while the latter three
    represent kinetic- and interaction terms for the nuclei(with the last one
    being the nuclei-particle interaction). The constant $M_k$ is the mass of
    nucleon $k$ and $Z_k$ is the corresponding atomic number.

    We assume the nuclei to be much heavier than the identical particles,
    meaning they move much slower, at which the system can be viewed as
    electrons moving around the vicinity of stationary nuclei.  This means the
    kinetic term for the nuclei vanish and the nuclei-nuclei interaction
    becomes a constant\footnote{Adding a constant to an operator does not alter
    the eigenvector, only the eigenvalues by the constant
    factor\cite{linalgDavid}.}. The approximation we end up with is the
    so-called \txtit{Born-Oppenheimer approximation} and the Hamiltonian is now
        \begin{equation}
            \mathcal{H} = \mathcal{H}_0 + \mathcal{H}_I
        \end{equation}
    where we have split the Hamiltonian in a \txtit{one-body} part and a
    \txtit{two-body} or \txtit{interaction} parts defined as
        \begin{equation}
            \mathcal{H}_0 \equiv - \frac{1}{2} \sum\limits_i \nabla^2_i +
            V\left(\blds{R},\blds{r}\right)
        \end{equation}
    and
        \begin{equation}
            \mathcal{H}_I \equiv \sum_{i<j} f\left(\blds{r}_i,
            \blds{r}_j\right)
        \end{equation}

\section{Slater Determinant and Permanent}
    Throughout \Arf{sec:3.1} we only referred to the wavefunction $\Psi$ as a
    state, a function closely connected to the probabilistic nature of the
    quantum particles. However, we have not given it a form. One possible
    solution is the \txtit{Hartree product} $\Psi_{\text{H}}$ defined as
        \begin{equation}
            \Psi_{\text{H}} = \prod_i\psi_i(\blds{r}_i)
        \end{equation}
    with $\{\psi\}_{i=0}^N$ being the orbitals which solve the single-particle
    Schrödinger equation for $H_0$. The Hartree-product is unfortunately a poor
    choice since it does not solve the $H_I$ part meaning it is not a
    physically valid solution. This comes from the fact that the
    Hartree-product does not take into account the fact that the particles in
    question are \txtit{identical particles}. Since the particles are
    identical, switching the labels on the particles shouldn't change the
    expectation value of some observable. If we run this remark through we
    end up with the conclusion that the state $\ket{\Psi}$ must be either
    symmetric or antisymmetric with the symmetric part being the \txtit{bosonic
    state} and antisymmetric being the \txtit{fermionic state}. The connection
    between antisymmetric states and fermions is called the \txtit{Pauli
    exclusion principle}.

    The problem with the Hartree-product is, with the above sentiment, that it
    is not symmetric nor antisymmetric. However we can transform it with an
    operator
        \begin{equation}
            \mathcal{B} \equiv \frac{1}{N!}\sum_P\sigma_b\mathcal{P}
        \end{equation}
    where $\sigma_b$ is defined as 
        \begin{equation}
            \sigma_b \equiv  
                \begin{cases}
                    1 \indent &\text{$b$ represents bosonic system} \\
                    (-1)^p \indent &\text{$b$ represents fermionic system}
                \end{cases}
        \end{equation}
    $\mathcal{P}$ is a permutation operator that switches the labels on particles
    \footnote{$P_{ij}\Psi_(\blds{r}_1,\dots,\blds{r}_i,\dots,\blds{r}_j,\dot,\blds{r}_N)
    = \Psi_(\blds{r}_1,\dots,\blds{r}_j,\dots,\blds{r}_i,\dot,\blds{r}_N)$
    \cite{compphysThijssen}.} and $p$ is the parity of permutations. The
    operator $\mathcal{B}$ has the following properties
        \begin{itemize}
            \item Applying $\mathcal{B}$ to itself doesn't change the operator
                meaning $\mathcal{B}^2 = \mathcal{B}$.
            \item The Hamiltonian $\mathcal{H}$ and $\mathcal{B}$
                \txtit{commute}, that is $\left[\mathcal{B},\mathcal{H}\right]
                = \left[\mathcal{H},\mathcal{B}\right]$.
            \item $\mathcal{B}$ is \txtit{unitary}, which means
                $\mathcal{B}^{\dagger}\mathcal{B}=\mathcal{I}$.
        \end{itemize}
    The solution $\Psi_T$ to the Schrödinger equation can now be written as
        \begin{equation}
            \Psi_T\left(\blds{r}\right) =
            \sqrt{N!}\mathcal{B}\Psi_{\text{H}}\left(\blds{r}\right)
            \label{eq:psiOpB}
        \end{equation}
    The antisymmetric case of $\mathcal{B}$ results in a \txtit{Slater
    determinant} 
        \begin{equation}
            \Psi^{\text{AS}}_T =
            \frac{1}{\sqrt{N!}}\sum_{P}(-1)^p\mathcal{P}_P\prod_i\psi_i
        \end{equation}
    while the symmetric case gives the so-called \txtit{permanent}\footnote{The
    permanent is basically just a determinant with all the negative signs
    replaced by positive ones.}.
        \begin{equation}
            \Psi^{\text{S}}_T =
            \sqrt{\frac{\prod\limits^N_{i=1}n_i!}{N!}}\sum_{P}\mathcal{P}_P\prod_i\psi_i
        \end{equation}
    Notice that the coordinated ${\blds{r}}_{i=0}^{N}$ is a bit of a sloppy
    notation as it also implicitly includes the spin orbitals. The extra factor
    in the symmetric case is to preserve the normalization due to number of
    particles, the $n_i$ factor is determined by the the quantum-number for
    state $i$.

\section{Variational Principle\label{sec:varPrinc}}
    One important remark is that the Slater determinant and the permanent do
    not solve the interaction part, but only serves as a so-called
    \txtit{ansatz} or guess on the true ground-state wavefunction. This is
    quite useful due to the \txtit{variational principle}. \\
    The Variational principle states that for any normalized function $\Psi$ in
    Hilbert Space\cite{GriffQuan} with a Hermitian operator $\mathcal{H}$ the
    minimum eigenvalue $E_0$ for $\blds{H}$ has an upper-bound given by the
    expectation value of $\mathcal{H}$ in the function $\Psi$. That is
        \begin{equation}
            E_0 \leq \ecp{\mathcal{H}} =
            \frac{\Braket{\Psi|\mathcal{H}|\Psi}}{\Braket{\Psi|\Psi}} 
            \label{eq:varPrinc}
        \end{equation}
    See \cite{GriffQuan} for proof and more.

    The mentioned ansatz is thereby guaranteed to give energies larger than or
    equal the true ground state energy meaning a minimization method is
    sufficient in order to get closer to this minimum. 
   
\section{Energy Functional\label{sec:energyFunc}}
    We can find a more convenient expression for this energy by using
    \Arf{eq:psiOpB} and \Arf{eq:varPrinc}. This gives us
        \begin{equation}
            E\left[\Psi\right] =
            N!\bra{\Psi_{\text{H}}}\mathcal{H}\mathcal{B}\ket{\Psi_{\text{H}}}
        \end{equation}
    where the hermitian and unitary property of $\mathcal{B}$ as well as the
    fact that $\mathcal{B}$ and $H$ commute have been used. This energy
    functional(functional in the sense that it is dependant on the wave
    function). Applying the $\mathcal{B}$ operator to the Hartree-product,
    pulling the sum out of the integrals and relabeling with the following
    definitions
        \begin{equation}
            \begin{aligned}
                \bra{p}h\ket{q} &\equiv
                \bra{\psi_p(\blds{r})}h(\blds{r})\ket{\psi_q(\blds{r})} =
                \int\psi^{*}_p(x)h(\blds{r})\psi_q(\blds{r})\md r \\
                \bra{pq}f\ket{rs} &\equiv
                \bra{\psi_p(\blds{r}_1)\psi_q(\blds{r}_2)}
                f(\blds{r}_1,\blds{r}_2)
                \ket{\psi_r(\blds{r}_1)\psi_s(\blds{r}_2)} \nonumber \\
                &= \int \psi_p(\blds{r}_1)\psi_q(\blds{r}_2)
                f(\blds{r}_1,\blds{r}_2) \psi_r(\blds{r}_1)\psi_s(\blds{r}_2)
                \md r
            \end{aligned}
        \end{equation}
    yields in\footnote{Notice also that $f_{12}$ is to imply integrals over two
    labels $r_1$ and $r_2$.}
        \begin{equation}
            E\left[\Psi\right] = \bra{p}H_0\ket{p} +
            \frac{1}{2}\sum_{p,q}\left[\bra{pq}f_{12}\ket{pq} \pm
            \bra{pq}f_{12}\ket{qp}\right]
            \label{eq:ch3Efunc}
        \end{equation}
    The first part is written with the assumption that the single-particle wave
    functions $\{\psi\}$ are orthogonal and the $1/2$ factor in front of the
    so-called \txtit{direct} and \txtit{exchange} terms\footnote{The direct
    term is just due to inherent behaviour of the charge of the particles
    (known as the Coulomb repulsion). The exchange term is a direct consequence
    of the probabilistic nature of the identical particles.} is due to the fact
    that we count the permutations twice in the sum when applying the
    $\mathcal{B}$ operator. The sign in the interaction term are chosen as
    positive for bosonic systems and negative for fermionic systems.

    The expression given in \Arf{eq:ch3Efunc} is the functional form we will
    use to derive the Hartree-fock equations in the following section.

\section{Hartree-Fock Theory\label{sec:HFtheory}}
    Hartree-Fock method is a many-body method for approximating the
    wavefunction of a stationary many-body quantum state and thereby also
    obtain an estimate for the energy in this state. \\
    The main idea is to represent the system as a \txtit{closed-shell} system
    and then variationally optimize the Slater determinant
    \cite{HelgakerMolElcTheory} and then iteratively solve the arising
    non-linear eigenvalue problem. This approach is not generally feasable as
    the size of the system increases, however for a closed-shell system in the
    quantum dot case it should be enough to build a basis, which is the goal
    here.
    For more details around the motivation and usages of the Hartree-Fock
    method see\cite{HelgakerMolElcTheory}. \\
    
    In this section we will derive the Hartree-Fock equations , following
    closely the literature by J.M Thjissen\cite{compphysThijssen}. 

    \subsection{Assumptions}
        Hartree-Fock method makes the following assumptions of the system
            \begin{itemize}
                \item \txtit{The Born-Oppenheimer approximation} holds. 
                \item All relativistic effects are negligible.
                \item The wavefunction can be described by a single
                    \txtit{Slater determinant}.
                \item The \txtit{Mean Field Approximation} holds.
            \end{itemize}
        With these inherent approximations the last one is the most important
        to take into account as it can cause large deviations from test
        solutions (analytic solutions, experimental data etc.) since the
        electron correlations is in reality, for many cases, not negligible.
        There exists many methods that try to fix this problem, but the
        \txtit{Variational Monte Carlo} (or VMC) is the method for deeper
        explorations in this Thesis, see \Arf{sec:QMC} for more details.
    
    \subsection{The $\blds{\mathcal{J}}$ and $\blds{\mathcal{K}}$ Operators}
        Before we begin with the Hartree-Fock equations it is desirable to
        rewrite the energy function obtained in \Arf{sec:energyFunc} (form
        given in \Arf{eq:ch3Efunc}) with two operators $\mathcal{J}$ and
        $\mathcal{K}$ defined as
            \begin{equation}
                \begin{aligned}
                    \mathcal{J} &\equiv \sum_k
                    \bra{\psi^{*}_k}f_{12}\ket{\psi_k} = \int
                    \psi^{*}_k(\blds{r})f_{12}\psi_k(\blds{r})\md r \\
                    \mathcal{K} &\equiv \sum_k
                    \bra{\psi^{*}_k}f_{12}\ket{\psi} = \int
                    \psi^{*}_k(\blds{r})f_{12}\psi(\blds{r})\md r \\
                \end{aligned}
            \end{equation}
        The $\mathcal{J}$ operator just gives the simple interaction-term while
        the $\mathcal{K}$ operator gives the exchange term with the arbitrary
        (notice no index) $\psi(\blds{r})$. The energy functional is thus
        rewritten to
            \begin{equation}
                E\left[\Psi\right] = \sum_i\Braket{\psi_i | h +
                \frac{1}{2}\left(\mathcal{J} \pm \mathcal{K}\right) | \psi_i}
                \label{eq:JKEfunc}
            \end{equation}
        where the one-body Hamiltonian is split into a sum of single particle
        functions as $H_0 = \sum\limits_i h(\blds{r}_i)$.

    \subsection{Hartree-Fock Equations}
        As a reminder. The wavefunctions $\{\psi\}$ in \Arf{eq:JKEfunc} are
        spin-orbitals with both a spacial part and a spin part. In order to
        obtain the Hartree-Fock equations we try to minimize the energy
        functional in order to obtain the ground-state energy for a many-body
        system. This is done by a variational method.

        The first observation to notice is the fact that variations in the
        spin-orbitals $\{\psi\}$ need to respect the spin-orthogonality
        relation
            \begin{equation}
                \braket{\psi_i}{\psi_j} = \delta_{ij}
            \end{equation}
        with $\delta_{ij}$ being the well-known Kronecker-delta. This property
        is essentially a constraint to the minimization problem and the method
        to be used is the \txtit{Lagrange multiplier method}\cite{linalgDavid},
        with the following \txtit{Lagrangian}
            \begin{equation}
                \mathcal{L} = \delta E\left[\Psi\right] -
                \sum_{ij}\Lambda_{ij}\left[\braket{\psi_i}{\psi_j} -
                \delta_{ij}\right]
            \end{equation}
        We know then that the minimum is reached when a displacement on the
        spin-orbitals $\psi_i\rarr\psi_i+\delta \psi_i$ results in an energy
        variation of zero meaning $\delta E\left[\Psi\right]=0$ in the minimum.
        Which gives the variational problem 
            \begin{equation}
                \sum_i \Braket{\delta\psi_i | h + \mathcal{J} \pm \mathcal{K} |
                \psi_i} - \sum_{ij} \Lambda_{ij}\braket{\delta\psi_i}{\psi_j} +
                \text{c.c} = 0
            \end{equation}
        where $\text{c.c}$ is a notation for the complex conjugate of the
        inner-products on its left-hand side.

        The shift in the spin orbitals $\{\delta \psi\}$ is arbitrary and the
        constraints are
        symmetric\footnote{$\braket{\psi_i}{\psi_j}=\braket{\psi_j}{\psi_i}^{*}
        \Rarr \Lambda_{ij}=\Lambda^{*}_{ji}$} meaning we can with the
        \txtit{Fock-operator}
            \begin{equation}
                \mathcal{F} \equiv h + \mathcal{J} \pm \mathcal{K}
            \end{equation}
        define the following eigenvalue problem
            \begin{equation}
                \mathcal{F}\psi_i = \sum_j\Lambda_{ij}\psi_j
            \end{equation}
        Choosing the Lagrange parameter $\Lambda_{ij}$ such that
        $\{\psi\}^N_{k=1}$ forms an orthonormal set for $\mathcal{F}$ with
        eigenvalues $\{\veps\}^N_{k=1}$. This reduces the eigenvalue equation
        to
            \begin{equation}
                \mathcal{F}\ket{\psi} = \blds{\veps}\ket{\psi}
                \label{eq:HFequations}
            \end{equation}
        with $\blds{\veps}=(\veps_0,\dots,\veps_N)$ being the set of
        eigenvalues of $\mathcal{F}$ meaning we have $N+1$ equations to be
        solved.

        If we only take the $N$ lowest eigenfunctions into the Slater the
        corresponding eigenenergy is referred to as the \txtit{Hartree-Fock
        energy} and is the estimated ground-state energy which the Hartree-Fock
        method gives. We can rewrite the energy functional with the
        eigenenergies to
            \begin{equation}
                E\left[\Psi\right] = \sum_i\Braket{\psi_i | \veps_i -
                \frac{1}{2}\left(\mathcal{J} \pm \mathcal{K}\right) | \psi_i}
                \label{eq:energyJKHF}
            \end{equation}
        In the derivation of the Hartree-Fock equations we only worked with
        spin-orbital functions $\{\psi\}$. However it is much more convenient
        to rewrite these in terms of spatial orbitals $\{\phi\}$ and integrate
        the spin-dependant part out. There are two ways of doing this and the
        two different approaches give the so-called \txtit{restricted
        Hartree-Fock} and \txtit{unrestricted Hartree-Fock} methods.

    \section{Restricted Hartree-Fock and Roothan-Hall-Equations}
        The restricted spin-orbitals are paired as\footnote{This is specialised
        for a two-spin system. For a system with more spin-states one needs to
        either choose different spacial-orbitals or add more such orbitals
        which effectively changes the energy-levels.}
            \begin{equation}
                \{\psi_{2l-1}, \psi_{2l}\} =
                \{\phi_l(\blds{r})\alpha(s),\phi_l(\blds{r})\beta(s)\}
                \label{eq:spinpairingHF}
            \end{equation}
        with $\alpha(s)$ and $\beta(s)$ being different spin-states (up and
        down). This pairing of spin-states with same and same spacial-orbitals
        means we can pull the spin degrees of freedom out from the
        $\mathcal{J}$ and $\mathcal{K}$ operators, reduce the sum to only run
        over half the states and multiply the entire sum by $2$. The result is
        that the restricted energy-functional reads
            \begin{equation}
                E\left[\Psi\right] = \sum\limits^N_{i=1}\veps_i -
                \sum\limits^{\frac{N}{2}}_{i=1} \Braket{i | 2\mathcal{J}
                \pm \mathcal{K} | i}
            \end{equation}
        Notice that the $\mathcal{K}$ operators sum only runs up to half the
        number of states.
        
        As the title suggests we are going to end up with a set of equations
        referred to as the \txtit{Roothan-Hall-equations}. We start by first
        expanding the spacial part $\{\phi\}$ of the spin orbitals $\{\psi\}$
        in some known orthonormal basis $\{\chi\}^L_{i=1}$
            \begin{equation}
                \phi_i(\blds{r}) = \sum\limits^L_{p=1} C_{pi}\chi_p(\blds{r})
            \end{equation}
        and introduce the \txtit{Fock-matrix} $F$(associated with the
        Fock-operator) with elements
            \begin{equation}
                F_{pq} = h_{pq} + \sum_{pq}\rho_{pq}\left(2D_{prqs} \pm
                D_{prsq}\right)
                \label{eq:FockRestrictedDef}
            \end{equation}
        We have here introduced a one-body matrix defined as
            \begin{equation}
                h_{pq} \equiv \Braket{p | h | q}
            \end{equation}
        a \txtit{density matrix} defined
        as\footnote{This is just the matrix formed by \begin{equation}
        \sum_i\ket{\phi_i}\bra{\psi_i}\end{equation} which is in quantum
        mechanics defined as the so-called \txtit{density matrix}.}
            \begin{equation}
                \rho_{pq} \equiv \sum\limits^{\frac{N}{2}}_{i=1}
                C_{pi}C^{*}_{qi}
                \label{eq:densitymatrixdef}
            \end{equation}
        and an interaction-matrix $D$ with elements
            \begin{equation}
                D_{pqrs} \equiv \Braket{pq | f_{12} | rs}
            \end{equation}
        for convenience. The implicit relabeling of $\chi_p(\blds{r})\rarr p$
        is also present in the above expression for the Fock-matrix. The
        Hartree-Fock equations (\Arf{eq:HFequations}) are then for the
        restricted case written as
            \begin{equation}
                \blds{F}\blds{C}_i = \blds{\veps}S\blds{C}_i
                \label{eq:eigvalrestr}
            \end{equation}
        with $S$ being the overlap matrix with elements
            \begin{equation}
                S_{pq} \equiv \Braket{p | q}
            \end{equation}
        This concludes the essential parts of the derivations of the
        Hartree-Fock equations. We also present the unrestricted case in
        \Arf{sec:UHFPNE} below. 
    
\section{Unrestricted Hartree-Fock and Poople-Nesbet-Equations\label{sec:UHFPNE}}
    In the spirit of completion we also write out the equations for the
    unrestricted case. The equations are in this case called the
    \txtit{Poople-Nesbet equations}. The derivations are exactly the same as
    for the restricted case, but without the spin-pairing in
    \Arf{eq:spinpairingHF}. We get two equations, one for the spin-up states
    and one for the spin-down\cite{compphysThijssen}. The equations
    are\footnote{These are just two Roothan-Hall equations.}
        \begin{equation}
            \begin{aligned}
                \blds{F}^{+}\blds{C}^{+} &=
                \blds{\veps}^{\+}\blds{S}\blds{C}^{+} \\
                \blds{F}^{-}\blds{C}^{-} &=
                \blds{\veps}^{-}\blds{S}\blds{C}^{-}
            \end{aligned}
        \end{equation}
    The elements of the Fock-matrices for spin-up and spin-down are
        \begin{equation}
            F^{\pm}_{pq} = h_{pq} + \sumll{k_{\pm}}\sumll{rs}
            C^{\pm\dagger}_{rk_{\pm}} C^{\pm\dagger}_{sk_{\pm}} \left[D_{prqs}
            - D_{prsq}\right] + \sumll{k_{\mp}}\sumll{rs}
            C^{\mp\dagger}_{rk_{\mp}} C^{\mp\dagger}_{sk_{\mp}} D_{prqs}
            \label{eq:UHFFockdef}
        \end{equation}
    The Hartree-Fock algorithm thus involves two eigenvalue problems at each
    iteration. Notice also that the summation index $k_{\pm}$ runs over the
    spin-up or spin-down states respectively and the $r$ and $s$ runs over all
    the spacial basis functions.

\section{Convergence and Mixing}
    The Hartree-Fock equations themselves are what we call a system of
    non-linear equations meaning that they need to be solved
    iteratively\footnote{and pray for convergence...}. A simple brute-force way
    of setting a convergence threshold is to say
        \begin{equation}
            \abs{\blds{\varepsilon}^{\text{HF}}_{\text{new}} -
            \blds{\varepsilon}^{\text{HF}}_{\text{old}}} < \text{threshold}
            \label{eq:HFconvthresh}
        \end{equation}
    The factors determining the achievement that is correct convergence to a
    ground-state is the basis set of choice, but even with a good set the
    notorious divergence can occur.  There are two methods in particular which
    are made to account for this. The first one is a simple
    \txtit{mixing}\cite{compphysThijssen} of the current density matrix and the
    one obtained at previous iteration
        \begin{equation}
            \blds{\rho}_{\text{new}} = \alpha\blds{\rho}_{\text{new}} + (1 -
            \alpha)\blds{\rho}_{\text{old}},\indent 0<\alpha\leq 1
            \label{eq:mixing}
        \end{equation}
    This method stems from the observation that for some systems the
    Hartree-Fock energies start to oscillate and the above mixing seems to be
    quite efficient at reducing the oscillations. \\
    The more popular approach however is the DIIS procedure, also known as
    Pulay mixing\cite{PULAY,PULAY2}. \\
    We use the simple mixing in this thesis.

    Here is also a figure describing the algorithm.
        \begin{figure}[H]
            \centering
            \begin{tikzpicture}[
                    >={Latex[width=2mm,length=2mm]},
                        base/.style = {rectangle, rounded corners, draw=black,
                                    minimum width=2cm, minimum height=0.5cm, text
                                    centered, font=\sffamily},
                        basecode/.style = {rectangle, rounded corners, draw=black,
                                    minimum width=2cm, minimum height=0.5cm, text
                                    centered, font=\sffamily, align=left},
                        activityStarts/.style = {base, fill=blue!30, drop shadow},
                        startstop/.style = {base, fill=red!30, drop shadow},
                        activityRuns/.style = {base, fill=green!30, drop shadow},
                        process/.style = {base, fill=white!15, font=\sffamily, drop shadow},
                        processcode/.style = {basecode, fill=white!15, font=\sffamily, drop shadow},
                    scale=0.8, 
                    node distance=1.5cm, 
                    every node/.style={fill=white, font=\sffamily},
                    align=center]
                \node (pre) [activityStarts] {
                    Pre-build two-body matrix $D_{pqrs}$ \\
                    Pre-build one-body matrix $h_{pq}$ \\
                    Pre-build overlap matrix $S_{pq}$ \\
                    Initialize $\blds{C}$
                };
                \node (Fock) [activityRuns, below of=pre, yshift=-1.0cm] {
                    Set Fock-matrix $\blds{F}$
                };
                \node (RHF) [process, above right of=Fock, xshift=3.0cm, yshift=-0.3cm] {
                    Use \Arf{eq:FockRestrictedDef} if RHF
                };
                \node (UHF) [process, below right of=Fock, xshift=3.0cm, yshift=0.1cm] {
                    Use \Arf{eq:UHFFockdef} if UHF
                };
                \draw[-] (Fock) to [out=15, in=180] (RHF);
                \draw[-] (Fock) to [out=-15, in=180] (UHF);
                \draw[->] (pre) -- (Fock);
                \node (eigcomp) [activityRuns, below of=Fock, yshift=-1.0cm] {
                    Solve generalized \\ 
                    eigenvalue problem. \\
                    $\blds{F}\blds{C} = \blds{\veps}S\blds{C}$
                };
                \draw[->] (Fock) -- (eigcomp);
                \node (density) [activityRuns, below of=eigcomp] {
                   Set density matrix $\blds{\rho}$.
                };
                \draw[->] (eigcomp) -- (density);
                \node (densityeq) [process, right of=density, xshift=2.5cm] {
                    \Arf{eq:densitymatrixdef}
                };
                \draw[-] (density) -- (densityeq);
                \node (mix) [activityRuns, below of=density] {
                    Perform mixing of $\blds{\rho}$.
                };
                \draw[->] (density) -- (mix);
                \node (mixeq) [process, right of=mix, xshift=2.5cm] {
                    \Arf{eq:mixing}
                };
                \draw[-] (mix) -- (mixeq);
                \node (conv) [startstop, below of=mix] {
                    Check convergence.
                };
                \node (conveq) [process, right of=conv, xshift=2.5cm] {
                    \Arf{eq:HFconvthresh}
                };
                \draw[->] (mix) -- (conv);
                \draw[-] (conv) -- (conveq);
                \node (no) [above left of=conv, xshift=-4.0cm] {
                    $\abs{\blds{\varepsilon}^{\text{HF}}_{\text{new}} -
                    \blds{\varepsilon}^{\text{HF}}_{\text{old}}} \geq \epsilon$
                };
                \draw[->] (conv) to [out=180,in=-90] (no);
                \node (keepOld) [process, above of=no, yshift=1.5cm, xshift=-0.5cm] {
                    Save energy eigenvalues and \\
                    density matrix(mixed). \\
                    $\blds{\varepsilon}^{\text{HF}}_{\text{old}} =
                    \blds{\varepsilon}^{\text{HF}}_{\text{new}}$ \\
                    $\blds{\rho}_{\text{old}} = \blds{\rho}_{\text{new}}$
                };
                \draw[->] (no) to [out=90,in=-90] (keepOld);
                \draw[->] (keepOld) to [out=90,in=180] (Fock);
                \node (yes) [below of=conv] {
                    $\abs{\blds{\varepsilon}^{\text{HF}}_{\text{new}} -
                    \blds{\varepsilon}^{\text{HF}}_{\text{old}}} < \epsilon$
                };
                \draw[->] (conv) -- (yes);
                \node (end) [activityStarts, below of=yes] {
                    Output energy
                };
                \node (endeq) [process, right of=end, xshift=3.0cm] {
                    \Arf{eq:energyJKHF, eq:ch3Efunc}
                };
                \draw[->] (yes) -- (end);
                \draw[-] (end) -- (endeq);
            \end{tikzpicture}
            \caption{Hartree-Fock Algorithm.}
        \end{figure}

\section{Quantum Monte Carlo\label{sec:QMC}}
    Quantum Monte Carlo, or QMC is a method for solving Schrödinger's equation
    by a statistical approach using so-called \txtit{Markov Chain} simulations
    (also called random walk). The nature of the wave function at hand is
    fundamentally a statistical model defined on a large configuration space
    with small areas of densities. The Monte Carlo method is perfect for
    solving such a system because of the non-homogeneous distribution of
    calculation across the space. An standard approach with equal distribution
    of calculation would then be a waste of computation time. 

    We will in this chapter address the Metropolis algorithm which is used to
    create a Markov chain and derive the equations used in the variational
    method.

    The chapter will use \txtit{Dirac Notation} \cite{GriffQuan} and all
    equations stated assume atomic units ($\hbar=m_e=e=4\pi\veps_0$) >> REF
    HERE ATOMIC UNITS <<.

    \subsection{The Variational Principle and Expectation Value of Energy}
        Given a Hamiltonian $\Ham$ and a trial wave function $\psiT$, the
        variational principle \cite{GriffQuan, NeOr} states that the
        expectation value of $\Ham$
            \begin{equation}
                E[\psi_T] = \ecp{\Ham} =
                \frac{\dinner{\psi_T}{\Ham}}{\pinner}
                \label{eq:ecpE}
            \end{equation}
        is an upper bound to the ground state energy
            \begin{equation}
                E_0 \leq \ecp{\Ham}
                \label{eq:ecpEBound}
            \end{equation}
        Now we can define our PDF as(see \Arf{susec:diffTHpdf} for a more
        detailed reasoning)
            \begin{equation}
                P(\mb{R}) \equiv \frac{\abs{\psi_T}^2}{\pinner}
                \label{eq:PDFdef}
            \end{equation}
        and with a new quantity
            \begin{equation}
                E_L(\mb{R};\mb{\alpha}) \equiv \frac{1}{\psiT}\Ham\psiT
                \label{eq:ELdef}
            \end{equation}
        the so-called local energy, we can rewrite \Arf{eq:ecpE} as
            \begin{equation}
                E[\psiT] = \ecp{E_L}
            \end{equation}
        The idea now is to find the lowest possible energy by varying a set of
        parameters $\mb{\alpha}$. This is done by numerical
        minimization(see \Arf{chapter:5}). We essentially minimize the
        expectation value of the energy were the expectation value itself is
        found with the Metropolis algorithm, see \Arf{susec:MHAlg}.

        An important property of the local energy is when we differentiate it
        with respect to one of the variational parameters $\{\alpha\}$ within
        the context of an expectation value. The result in this case would be
        zero. This is easily seen by direct calculation
            \begin{align}
                \ecp{\prd{\alpha}[E_L]} &= \int \frac{\abs{\psi}^2\prd{\alpha}
                \left[\frac{1}{\psi}H\psi\right]}{\int \abs{\psi}^2 \md r} \md
                r \nonumber \\
                &= \int \frac{\abs{\psi}^2 \frac{\psi^{*}\prd{\alpha}(H\psi) -
                (H\psi^{*}) \prd{\alpha}[\psi]}{\abs{\psi}^2}}{\int
                \abs{\psi}^2 \md r} \md r \nonumber \\
                &= \int \frac{\psi^{*}H\prd{\alpha}[\psi] -
                \psi^{*}H\prd{\alpha}[\psi]}{\int \abs{\psi}^2 \md r} \md r
                \nonumber \\
                &= 0
                \label{eq:expELdelzero}
            \end{align}
        We have used the fact that $H$ is not dependant on any variational
        parameter and used the hermitian properties\cite{GriffQuan} of $H$ to
        justify the movement of $H$ within the integral.  

        This neat result presented in \Arf{eq:expELdelzero} will show its
        usefulness in the minimization when derivatives of the expectation
        value come into play since finding the derivative of the local energy
        would be much more of a hassle.\footnote{The differentiation of the
        wavefunction is enough to reduce the quality of life on its own!} \\
        We also write the derivative with respect to a variational parameters
        $\alpha$ of the expectation value $\ecp{E}$ here for reference
            \begin{equation}
               \prd{\alpha}[\ecp{E}] =
                2\left(\ecp{\frac{E_L}{\psi}\prd{\alpha}[\psi]} -
                \ecp{E}\ecp{\frac{1}{\psi}\prd{\alpha}[\psi]}\right)
            \end{equation}
        We have also applied \Arf{eq:expELdelzero} here.

    \subsection{Metropolis-Hastings Algorithm\label{susec:MHAlg}}
        The Metropolis algorithm bases itself on moves (also called
        transitions) as given in a Markov process\cite{NeOr, FeWa}. This
        process is given by
            \begin{equation}
                w_i(t+\veps) = \sum_j\ufij{w}{i}{j}w_j(t)
            \end{equation}
        where $w(j\rarr i)$ is just a transition from state $j$ to state $i$.
        In order for the transition chain to reach a desired convergence while
        reversibility is kept, the well known condition for detailed balance
        must be fulfilled\cite{statmech}. If detailed balance is true, then the
        following relation is true
            \begin{equation}
                w_i \ufij{T}{i}{j}\ufij{A}{i}{j} = w_j \ufij{T}{j}{i}\ufij{A}{j}{i}
                \Rarr \frac{w_i}{w_j} =
                \frac{\ufij{T}{j}{i}\ufij{A}{j}{i}}{\ufij{T}{i}{j}\ufij{A}{i}{j}}
                \label{eq:detailedBalance}
            \end{equation}
        We have here introduced two scenarios, the transition from
        configuration $i$ to configuration $j$ and the reverse process $j$ to
        $i$. Solving the acceptance $A$ for the two cases where the ratio in
        \ref{eq:detailedBalance} is either $1$(in which case the proposed state
        $j$ is accepted and transitions is made) and when the ratio is less
        then $1$. The Metropolis algorithm would in this case not automatically
        reject the latter case, but rather reject it with a proposed uniform
        probability. Introducing now a probability distribution function(PDF) $P$
        the acceptance $A$ can be expressed as
            \begin{equation}
                \ufij{A}{i}{j} =
                \text{min}\left(\frac{\ufij{P}{i}{j}}{\ufij{P}{j}{i}}
                \frac{\ufij{T}{i}{j}}{\ufij{T}{j}{i}} ,1\right)
                \label{eq:metropolisAcceptance}
            \end{equation}
        The so-called selection probability $T$ is defined specifically for
        each problem. For our case the PDF in question is the absolute square
        of the wave function and the selection $T$ is a Green's function
        derived in \Arf{susec:impSamp}.  The algorithm itself would then be
        \begin{figure}[H]
            \centering
            \begin{tikzpicture}[
                >={Latex[width=2mm,length=2mm]},
                    base/.style = {rectangle, rounded corners, draw=black,
                                minimum width=2cm, minimum height=0.5cm, text
                                centered, font=\sffamily},
                    basecode/.style = {rectangle, rounded corners, draw=black,
                                minimum width=2cm, minimum height=0.5cm, text
                                centered, font=\sffamily, align=left},
                    activityStarts/.style = {base, fill=NavyBlue!30, drop shadow},
                    startstop/.style = {base, fill=Maroon!25, drop shadow},
                    startstopcode/.style = {basecode, fill=Maroon!25, drop shadow},
                    activityRuns/.style = {base, fill=ForestGreen!25, drop shadow},
                    process/.style = {base, fill=white!15, font=\sffamily, drop shadow},
                    processcode/.style = {basecode, fill=white!15, font=\sffamily, drop shadow},
                scale=0.8, 
                node distance=1.5cm, 
                every node/.style={fill=white, font=\sffamily},
                align=center]
                \node (1) [activityStarts] {
                    Pick initial state $i$ at random.
                };
                \node (2) [activityStarts, below of=1] {
                    Pick proposed state at random in accordance to
                    $\ufij{T}{j}{i}$.
                };
                \node (3) [activityStarts, below of=2] {
                    Accept state according to $\ufij{A}{j}{i}$.
                };
                \node (a) [activityRuns, below left of=3, xshift=-3cm, yshift=-0.1cm] {
                    Accepted \\
                    $j = i$
                };
                \node (r) [startstop, below right of=3, xshift=3cm, yshift=-0.1cm] {
                    Rejected \\
                    $i = j$
                };
                \draw[<-] (3) to [out=180,in=90] (a);
                \draw[<-] (3) to [out=0,in=90] (r);
                \node (4) [activityStarts, below of=3, yshift=-1cm] {
                    Number of states satisfied?
                };
                \node (yes) [below of=4] {
                    Yes
                };
                \node (no) [above left of=4, xshift=-5cm] {
                    No
                };
                \draw[->] (4) to [out=180, in=-90] (no);
                \draw[->] (no) to [out=90, in=180] (2);
                \node (5) [activityStarts, below of=yes] {
                    End.
                };
                \draw[->] (1) -- (2);
                \draw[->] (2) -- (3);
                \draw[->] (3) -- (4);
                \draw[->] (4) -- (yes);
                \draw[->] (yes) -- (5);
            \end{tikzpicture}
            \caption{Metropolis algorithm.}
        \end{figure}

    \subsection{Diffusion Theory and the PDF\label{susec:diffTHpdf}}
        The motivation for the use of diffusion theory is described very well
        in the Masters thesis of Jørgen Høgberget\cite{jorgenThesis}. The
        essential results were a Green's function propagator arising from the
        \txtit{short time approximation} and the observation of interest here
        which has to do with the statistics describing the expectation value,
        it states that \txtit{any} distribution may be applied in calculation,
        however if we take a close look at the local energy(\Arf{eq:ELdef}) we
        see that the local energy is not defined at the zeros of $\psiT$ for
        all distributions.  This means that an arbitrary PDF does not guarantee
        generation of points which makes $\psi_T=0$. This can be overcome by
        introducing the square of the wave function to be defined as the
        distribution function as given in \Arf{eq:PDFdef}. This basically means
        that when using Quantum Monte-Carlo methods, the incorporation of the
        fact that when an energy is more undefined(meaning $\psi_T\rarr 0$ the
        less probable that point is actually means the generation of states in
        which $\psi_T$ is removed. The next chapter will derive the method that
        includes this observation into play, the so-called \txtit{importance
        sampling}.

    \subsection{Importance Sampling\label{susec:impSamp}}
        Using the selection probability mentioned in \Arf{susec:MHAlg}, the
        Metropolis-Hastings algorithm is called an \txtit{Importance sampling}
        because it essentially makes the sampling more concentrated around
        areas where the PDF has large values. \\

        Because of the intrinsic statistical property of the wave function
        Quantum Mechanics can be modelled as a diffusion process, or more
        specifically, an \txtit{Isotropic Diffusion Process} which is
        essentially just a random walk model. Such a process is described by
        the Langevin equation with the corresponding Fokker-Planck equation
        describing the motion of the walkers(particles). See \cite{numstoch,
        compphysThijssen, vmc} for details. The full derivations presented here
        follows closely the descriptions in \cite{compphysThijssen, vmc}. Let
        us start off with the Langevin equation
            \begin{equation}
                \langevin
                \label{eq:langevin}
            \end{equation}
        and apply Euler's method and obtain the new positions
            \begin{equation}
                \rnew = \rold + D\Fold\Delta t + \xi
                \label{eq:rnewdef}
            \end{equation}
        with the $r$'s being the new and old positions in the Markov chain
        respectively and $\Fold=F(\rold)$. The quantity $D$ is a diffusion
        therm equal to $1/2$ due to the kinetic energy(remind of natural units)
        and $\xi$ is a Gaussian distributed random number with $0$ mean and
        $\sqrt{\Delta t}$ variance.

        As mentioned a particle is described by the Fokker-Planck equation
            \begin{equation}
                \FokkerPlanck
                \label{eq:FokkerPlanckDef}
            \end{equation}
        With $P$ being the PDF(in current case the selection probability) and
        $F$ being the drift therm. In order to achieve convergence, that is a
        stationary probability density, we need the left hand side to be zero
        in \Arf{eq:FokkerPlanckDef} giving the following equation
            \begin{equation}
                \prd{x_i}[P][2] = P\prd{x_i}[\mb{F_i}] + \mb{F_i}\prd{x_i}[P]
            \end{equation}
        with the drift-therm being on the form $\mb{F}=g(x)\prtl P/\prtl x$ we
        finally have that
            \begin{equation}
                \mb{F} = \frac{2}{\psi_T}\nabla \psi_T
                \label{eq:qForceDef}
            \end{equation}
        This is the so-called \txtit{Quantum Force} which pushes the walkers
        towards regions where the wave function is large. \\
        The missing part now is to model the selection probability in
        \Arf{eq:metropolisAcceptance}. Inserting the quantum force into the
        Focker-Planck equation(\Arf{eq:FokkerPlanckDef} the following diffusion
        equation appears
            \begin{equation}
                \prd{t}[P] = -D\nabla^2 P
                \label{eq:diffeqFP}
            \end{equation}
        Applying the \txtit{Fourier Transform} to spatial coordinate $r$ in
        \Arf{eq:diffeqFP}, the equation is transformed to
            \begin{equation}
                \prd{t}[P(\blds{s},t)] = -Ds^2 P(\blds{s},t)
            \end{equation}
        which has solution
            \begin{equation}
                P(\blds{s},\Delta t) = P(\blds{s},0)\me^{Ds^2\Delta t}
                \label{eq:Punfin}
            \end{equation}
        Now we need to find the constant $P(\blds{s},0)$, and as is apparent
        with $t=0$, we will make use of an initial condition. The initial
        positions are spread out from origin, that is $D\Delta t\blds{F}_j$. We
        can express this with a \txtit{Dirac-delta function}\cite{GriffQuan,
        boasMath} giving
            \begin{equation}
                P(s,0) = \delta\left(\blds{r}_i - D\Delta t\blds{F}_j\right)
            \end{equation}
        Inserting this into \Arf{eq:Punfin} and making the inverse Fourier
        transform yields the following Green's function as solution
            \begin{equation}
                P\left(a, b, \Delta t\right) = \frac{1}{\sqrt{4\pi D\Delta t}}
                \exp(-\frac{\left(\blds{r}_a - \blds{r}_b - D\Delta t
                \blds{F}_b\right)^2}{4D\Delta t})
            \end{equation}
        This expression is precisely the selection probability $T$, Notice also
        that the indices $a$ and $b$ label a state transition $a\rarr b$ and not
        particle indices. The full transition probability needs to be summed
        over for all particles since we only solved the Focker-Planck equation
        for $1$ particle(since the other solutions are found in the exact same
        manner). For clarity the full selection probability ratio is
            \begin{equation}
                \frac{T(b,a,\Delta t)}{T(a,b,\Delta t)} = \sum_i
                \exp(-\frac{\left(\blds{r}^{(b)}_i - \blds{r}^{(a)}_i - D\Delta
                t \blds{F}^{(a)}_i\right)^2}{4D\Delta t}
                +\frac{\left(\blds{r}^{(a)}_i - \blds{r}^{(b)}_i - D\Delta t
                \blds{F}^{(b)}_i\right)^2}{4D\Delta t})
            \end{equation}

    \subsection{The Trial Wavefunction: One-Body}
        The trial wave function is generally an arbitrary choice specific for
        the problem at hand, however it is in most cases favorable to expand
        the wave function in the eigenbasis (eigenstates) of the Hamiltonian
        since they form a complete set. This can be expressed as
            \begin{equation}
                \psiT = \sum_k C_i\psi_k(\mb{R};\mb{\alpha})
            \end{equation}
        where the $\psi_i$'s are the eigenstates of the Hamiltonian. The
        coefficients can be found by any method preferable and is the usual
        procedure is to use a set of basis functions and then minimize to find
        the coefficients $\{C\}_{k=1}^L$. We use the Hartree-Fock method to
        minimize in this thesis. The trial wavefunction is also generally
        expressed as a \txtit{Slater determinant} for the fermionic case and a
        general product for bosonic systems\cite{GriffQuan, NeOr, FeWa, DiVNe}
        We will explain the fermionic case shortly since it is the main focus
        here and since the bosonic wavefunction is simple to express. The
        Slater is expressed as
            \begin{equation}
                \phiT = \det(\blds{\Phi(\blds{R};\blds{\alpha})}) \xi(s)
            \end{equation}
        where the \txtit{Slater matrix} $\blds{\Phi}$ has elements
            \begin{equation}
                \Phi_{ij} = \phi_{n_j}(\blds{r}_i;\blds{\alpha})
            \end{equation}
        such that each row is evaluated for particle $i$ and each column is for
        a quantum number $n_j$ dependant on the basis used. The $\xi(s)$ is the
        spin-dependant part. Notice also that we switched the labeling from
        $\Psi_T$ to $\Phi_T$. This is to make a distinction between
        \txtit{one-body} and \txtit{correlation} terms. The latter will be
        introduced later in \Arf{susec:TWFJastrow}. In this case the
        \txtit{single-particle} functions $\phi_j(r)$ are expanded in some
        basis(i.e Hartree-Fock).

    \subsection{The Trial Wavefunction: Splitting the Slater Determinant}
        An important part of the trial-wavefunction presented here is that the
        one-body term is \txtit{independent} of spin, that is, the Hamiltonian
        is not explicitly dependant on the spin degrees of freedom. For the
        case of a Hamiltonian with an inherent spin part The following
        splitting of the Slater determinant is not valid! In that case the
        expectation value(presented in the variational principal in
        \Arf{eq:varPrinc}) would be a product of the expectation value over the
        spin-independent part of the Hamiltonian and the expectation value over
        the remaining spin-dependant parts\cite{spinDep}. The results presented
        here is however for systems of spin-independent systems, and in those
        cases the spin-part is essentially just another label which can be
        integrated out(similar to the procedure with the restricted
        Hartree-Fock method). For the splitting with a \txtit{spin-dependent}
        Hamiltonian see \cite{splitDet} and \cite{basicMB}. \\
        
        The procedure is simply to arrange the basis functions in such a way
        that we have $N/2$ single-particle functions in spin-up and the same
        basis functions for spin-down. This means that the Slater is
            \begin{equation}
                \frac{1}{N!} \begin{pmatrix}
                    \phi_1(\blds{r}_1)\xi_{\urw} & \dots &
                    \phi_{N/2}(\blds{r}_1)\xi_{\urw} &
                    \phi_1(\blds{r}_1)\xi_{\drw} & \dots &
                    \phi_{N/2}(\blds{r}_1)\xi_{\drw} \\
                    \vdots & \vdots & \vdots & \vdots & \vdots & \vdots \\
                    \phi_1(\blds{r}_N)\xi_{\urw} & \dots &
                    \phi_{N/2}(\blds{r}_N)\xi_{\urw} &
                    \phi_1(\blds{r}_N)\xi_{\drw} & \dots &
                    \phi_{N/2}(\blds{r}_N)\xi_{\drw} \\
                \end{pmatrix}
            \end{equation}
        This restructuring of the single-particle states implies that
            \begin{equation}
                \det(\blds{\Phi}) \propto \det(\blds{\Phi_{\urw}})
                \det(\blds{\Phi_{\drw}})
            \end{equation}
        where we have defined
            \begin{equation}
                \begin{aligned}
                    \blds{\Phi_{\urw}} &= \begin{pmatrix}
                        \phi_1(\blds{r}_1) & \dots & \phi_{N/2}(\blds{r}_1) \\
                        \vdots & \ddots & \vdots \\
                        \phi_1(\blds{r}_{N/2}) & \dots &
                        \phi_{N/2}(\blds{r}_{N/2})
                    \end{pmatrix} \\
                    \blds{\Phi_{\drw}} &= \begin{pmatrix}
                        \phi_1(\blds{r}_{N/2+1}) & \dots &
                        \phi_{N/2}(\blds{r}_{N/2+1}) \\
                        \vdots & \ddots & \vdots \\
                        \phi_1(\blds{r}_{N}) & \dots & \phi_{N}(\blds{r}_{N})
                    \end{pmatrix}
                \end{aligned}
            \end{equation}
        Which in this sense means to put the first $N/2$ particles in spin-up
        configuration and the remaining in spin-down and use the same
        single-particle functions.  On a technical note, this rewriting is an
        approximation. However it can be shown(again see \cite{splitDet}) that
        the expectation value is still the same.\\
        The one-body term is now rewritten to
            \begin{equation}
                \det(\blds{\Phi}) = \det(\blds{\Phi_{\urw}})
                \det(\blds{\Phi_{\drw}})
                \label{eq:slatersplitdef}
            \end{equation}

    \subsection{The Trial Wavefunction: Jastrow Factor\label{susec:TWFJastrow}}
        As mentioned, we model the wavefunction as a product of the one-body
        Slater and a correlation part known as a \txtit{Jastrow factor}. The
        Jastrow can have many forms, but is build to exert some key features.
        It should\cite{jastrow}. 
        \begin{itemize}
            \item Be dependent on the inter-particle distances.
            \item Approach unity at large distances.
            \item Vanish when particles are close to one another.
        \end{itemize}
        In this thesis we explore two forms, the popular
        \txtit{Pad\'e-function} and a more resent one build with a specific
        neural network known as a \txtit{Boltzmann Machine}.

    \subsubsection{Cusp Condition\label{sususec:cusp_condition}}
        The derivations for the cusp conditions is described well in the master
        thesis of Lars Eivind Lervåg\cite{larsEivindThesis}. We will here
        present the result of his for both two and three dimensions. Assuming
        the Jastrow factor $J$ has the form
            \begin{equation}
                J = \prod_{i<j} \me^{f\left(r_{ij}\right)}
            \end{equation}
        and considering the local energy as two particles $i$ and $j$ approach,
        the final results were the following conditions
            \begin{equation} 
                    \begin{aligned}
                        \prd{r_{ij}}[f] \Bigg|_{r_{ij}=0} &=
                        \frac{1}{D-1},&\indent& \text{anti-parallel spin} \\
                        \prd{r_{ij}}[f] \Bigg|_{r_{ij}=0} &=
                        \frac{1}{D+1},&\indent& \text{parallel spin}
                    \end{aligned}
            \end{equation}
        which are known as the cusp conditions.
    
    \subsubsection{Pad\'e Function\label{sususec:TWFPadeJastrow}} 
    A popular form is the \txtit{Pad\'e-Jastrow} function. It is defined
    as\cite{basicMB, MCAbInitoChem}
            \begin{equation}
                J_{\text{Pad\'e}} =
                \exp\left(\sum\limits_{i<j}\frac{\sum\limits_la^{(l)}_{ij}
                r^l_{ij}}{1 + \sum\limits_l \beta_l r^l_{ij}}\right)
            \end{equation}
        where the $\beta_l$'s are variational parameters. This function follows
        the key features mentioned for a correlation factor and the
        restrictions
            \begin{equation}
                a^{\text{2D}}_{ij} = \left\{\begin{aligned}
                        \frac{1}{3},&\indent& &\text{parallel spin} \\
                        1,&\indent& &\text{anti-parallel spin} \\
                    \end{aligned}\right.
            \end{equation}
            \begin{equation}
                a^{\text{3D}}_{ij} = \left\{\begin{aligned}
                        \frac{1}{4},&\indent& &\text{parallel spin} \\
                        \frac{1}{2},&\indent& &\text{anti-parallel spin} \\
                    \end{aligned}\right.
            \end{equation}
        on the factor $a$ along with the fact that $a^{(l)}_{kj}=0$ for $l>1$,
        means we may relabel $a^{(l)}_{ij}\rarr a_{ij}$ and $\beta_l\rarr
        \beta$ and reduce the Pad\'e-Jastrow factor to
            \begin{equation}
                J_{\text{Pad\'e}} = \exp\left(\sum\limits_{i<j}\frac{a_{ij}
                r_{ij}}{1 + \beta r_{ij}}\right)
                \label{eq:padereduceddef}
            \end{equation}
        For expressions concerning the gradient and Laplacian, see
        \Arf{appendix:B}.

    \subsubsection{Simple Exponential}
        >> REF THIS << With the desirable features for a correlation function
        in mind, one of the simplest forms for a correlation factor is a scaled
        exponential evaluated with the inter-particle distance, this
        is\footnote{The $a$'s can also be variational parameters.}
            \begin{equation}
                J = \exp(\sumll{i<j} a_{ij} r_{ij})
            \end{equation}
        The $a$ would be defined the same as with the Pade\'e function. In
        itself this function is fairly useless as it doesn't model correlations
        very well for different systems or even the same system with different
        parameters. However we will present a new type of function in the next
        section which does not give raise to any cusp conditions, but has a
        more flexible form and may in connection with this simple exponential
        give us all the necessary properties desired in a correlation function.

    \subsubsection{The Trial Wavefunction: NQS Wavefunction
    \label{sususec:NQSJastrow}}
        A more resent and completely different approach is to model the
        wavefunction with a \txtit{neural network} as presented by Carleo and
        Troyer\cite{CarleoANN}. The approach used here is based on the
        \txtit{Restricted Boltzmann Machine} as described by
        Hinton\cite{RBMpractical}. The form is
            \begin{equation}
                J_{\text{NQS}} = \exp(-\suml{i=1}{N} \frac{\left(\blds{r}_i -
                \blds{a}_i\right)^2}{2\sigma^2})\prod\limits^{M}_j \left(1 +
                \exp(b_j + \suml{i=1}{N}\suml{d=1}{D}
                \frac{x^{(d)}_iw_{i+d,j}}{\sigma^2})\right)
                \label{eq:nqsdef}
            \end{equation}
        The derivatives of this are given in \Arf{appendix:B}.

    \subsection{Connect the Jastrows}
        In the previous section we presented four functions that can be used to
        model correlations in quantum systems, however they all had some
        limitations as well as advantageous properties. The question then
        remains, can we create a function which exhibits all the nice
        properties mentioned and none of the limitations? The answer is a bit
        ambivalent. We don't really know for sure, but a good start is to
        multiply the functions having the cusp conditions in order (Pad\'e and
        exponential) with the NQS wavefunction and use this product as the new
        jastrow. The motivations for this is simple, we want the cusp
        conditions introduced by the old popular functions, but also want a
        more flexible type of function which can take care of other(possible)
        correlations in the system. How well this actually performed is
        presented in >> REF RESULTS WHEN YOU HAVE SOME (GO GET THEM!) <<. We
        will write out the exact form for clarity.
            \begin{equation}
                J = \exp(\sum\limits_{i<j}a_{ij} r_{ij})
                \exp(-\suml{i=1}{N} \frac{\left(\blds{r}_i -
                \blds{a}_i\right)^2}{2\sigma^2})\prod\limits^{M}_j \left(1 +
                \exp(b_j + \suml{i=1}{N}\suml{d=1}{D}
                \frac{x^{(d)}_iw_{i+d,j}}{\sigma^2})\right)
            \end{equation}
            \begin{equation}
                J = \exp\left(\sum\limits_{i<j}\frac{a_{ij} r_{ij}}{1 + \beta
                r_{ij}}\right) \exp(-\suml{i=1}{N} \frac{\left(\blds{r}_i -
                \blds{a}_i\right)^2}{2\sigma^2})\prod\limits^{M}_j \left(1 +
                \exp(b_j + \suml{i=1}{N}\suml{d=1}{D}
                \frac{x^{(d)}_iw_{i+d,j}}{\sigma^2})\right)
            \end{equation}
        It might not be completely clear as to why this function still makes it
        so that the total wavefunction satisfies the cusp conditions. The
        reason lies in the derivation of the conditions, the procedure in which
        the conditions presented in \Arf{sususec:cusp_condition} were found the
        only consideration was within the terms involving $r_{ij}$ in the local
        energy, all other terms were canceled out or were otherwise
        non-divergent meaning the product between the jastrow would yield the
        same conditions.

\section{Density Matrices}
    The Slater wavefunction presented throughout this chapter is quite
    complicated and the solution-space for the Schrödinger equation with said
    Slater is too large to extract any useful physical properties directly.
    Luckily we have one part of the physical system we may extract from a
    variational calculation, the so-called \txtit{one-body density}. From
    quantum mechanics the one-body density for a normalized wavefunction is
    defined as
        \begin{equation}
            \Lambda(\blds{r},\blds{r}') = N\int
            \psi^{*}(\blds{r},\blds{r}_2,\dots,\blds{r}_N)
            \psi(\blds{r}',\blds{r}_2,\dots,\blds{r}_N)\md\blds{r}_2,\dots,\md\blds{r}_N
        \end{equation}
    which is just to integrate the wavefunction over all spacial coordinates
    except $\blds{r}$ and $\blds{r}'$. The question now is what exactly does
    this mean, what information does this integral give us? The one-body density
    tells us that the element
        \begin{equation}
            N\int
            \psi^{*}(\blds{r})\psi(\blds{r})\md\blds{r}_2,\dots,\md\blds{r}_N =
            N\times P\left(\text{Finding a particle within volume } \md\blds{r}
            \text{around point } \blds{r}\right)
        \end{equation}
    with $P$ denoting a probability. This might not give any tingles at first
    when it comes to extraction of physical properties of the system, however
    one has to keep in mind that the actual \txtit{configuration} of particles
    within the spacial solution-grid gives a direct insight into the actual
    physics of the system. A good example of this is the distribution
        \begin{equation}
            N\int \psi^{*}(\blds{r}_1)\psi(\blds{r}_1)\md\blds{r}_{12} = 0
        \end{equation}
    This integral is a \txtit{two-body density} and essentially what we have
    asked ourselves now is, what is the probability of finding two particles in
    the same exact state? The answer is apparently \txtit{zero} and we see that
    the all-time famous Pauli-exclusion principle practically appears straight
    out of the density definition. \\
    With this result as motivation in mind, it is not unreasonable to question
    if density-matrices might possess more information about the physical
    nature of the system in question. \\
    Before we tackle the procedure of calculating the densities with the
    Metropolis algorithm, let us express the full $N$-body density. The
    expressions are directly copied from this insightful article by Per-Olov
    L\"owdin\cite{onebodydens}. Density matrix of order $p$ is defined to be
        \begin{equation}
            \Lambda^{(p)}\left(\blds{r}'_1,\dots,\blds{r}'_p|\blds{r}_1,\dots,\blds{r}_p\right)
            = {N\choose p} \int \psi^{*}(1',2',\dots,p',\dots,N)
            \psi(1,2,\dots,p,\dots,N) \md\blds{r}_{12}',\dots,\md\blds{r}_{p}'
        \end{equation}
    The integrals are over all permutations of $\blds{r}_{12}'$. \\
    In order to actually find this density with the Metropolis algorithm we
    need to rewrite it in the same manner as with the local energy, which is to
    introduce $\abs{\psi}^2$. For the fermionic case this rewriting gives us
    the following matrix elements
        \begin{equation}
            \Lambda_{ij} = N\int\phi^{*}_i(\blds{r}_1)\phi_j(\blds{r}')
            \frac{\psi(\blds{r}',\blds{r}_2,\dots,\blds{r}_N)}{\psi(\blds{r}_1,\dots,\blds{r}_N)}
            \abs{\psi(\blds{r}_1,\dots,\blds{r}_N)}^2
            \md\blds{r}'\md\blds{r}_1\dots\md\blds{r}_N
        \end{equation}
    where $\phi$ are the single-particle basis functions within the Slater
    $\psi$. \\
    The elements of the $\blds{\Lambda}$ matrix can then be calculated by
    simple counting. The basic premise is to essentially create a histogram of
    particle counts with the bins being various radial distances up to some
    cutoff $r_{\text{max}}$. Then at every Metropolis step (after the test) we
    increment each value in the bins array with the number of particles
    currently within each respective bin. The array is then normalized by the
    usual total sum (sum of all values in the bins array), but also the radial
    distance to the power $D$ to the right-most edge of each bin. The reason
    for this additional normalization is to account for the radial
    contributions to the configurations, a larger radial distance means the
    sparsity of the particle-density increases. We are however only interested
    in the non-dimensional count within each vicinity meaning we need to divide
    away the dimensional volume element of each bin. The element for a bin $n$
    is found to be\footnote{Constant proportional factors are dropped due to
    them being canceled in the histogram normalization(division of the sum of
    whole bin array).}
        \begin{align}
            V^{(D)}_n &= r^D_{n+1} - r^D_n \nonumber \\
            &= (r_n+\Delta r)^D - r^D_n
        \end{align}
    In order to keep this stable we also notice that if the number of bins is
    satisfyingly large higher orders of $\Delta r$ vanish and only the terms
    with linear factors of $\Delta r$ give a significant contribution. Notice
    also that the highest order $r^D_n$ gets canceled out. All this together
    gives
        \begin{equation}
            V^{(D)}_n = Dr_n^{D-1}\Delta r
        \end{equation}
    The factor $D$ in front can be dropped as the histogram normalization
    cancels it anyways. \\
    The resulting one-body densities are presented visually in >> REF THESE
    FIGURES (GO GET EM!) <<
