\begin{thebibliography}{}
    \bibitem{GriffQuan}
        David J. Griffiths,
        \textit{Introduction to Quantum Mechanics Second Edition},
        Pearson, 2005, ISBN 0-13-111892-7.
    \bibitem{calcVar}
        Bernard Dacorogna, Chapters 2 and 6,
        \textit{Introduction to the Calculus of Variations (3rd Edition)},
        World Scientific, 2014, ISBN 978-1-78326-551-0.
    \bibitem{calcVarSpring}
        Bruce Van Brunt,
        \textit{The Calculus of Variations},
        Springer, 2004, ISBN 0-387-40247-0.
    \bibitem{basicMB}
        Morten Hjort-Jensen,
        \textit{Computational Physics: Hartree-Fock methods and introduction to Many-Body Theory.} \\
        \url{https://www.github.com/CompPhysics/ComputationalPhysics2/blob/gh-pages/doc/pub/basicMB/pdf}, 2017
    \bibitem{vmc}
        Morten Hjort-Jensen,
        \textit{Computational Physics: Variational Monte Carlo methods.} \\
        \url{https://www.github.com/CompPhysics/ComputationalPhysics2/blob/gh-pages/doc/pub/vmc/pdf}, 2017
    \bibitem{taut}
        M. Taut, 
        \textit{Two Electrons in an External Oscillator Potential; Particular analytic solutions of a Coulomb Correlation Problem},
        Physical Review A 48, 3561, 1993.
    \bibitem{NeOr}
        Negele and Orland,
        \textit{Quantum Many-Particle Systems},
        Addison-Wesley.
    \bibitem{FeWa}
        Fetter and Walecka,
        \textit{Quantum Theory of Many-Particle Systems},
        Mcgraw-Hill, 1971.
    \bibitem{DiVNe}
        Dickhoss and Van Neck,
        \textit{Many-Body Theory Exposed},
        World Scientific, 2006
    \bibitem{shermorInv}
        Jack Sherman and Winifred J. Morrison,
        \textit{Adjustment of an Inverse Matrix Corresponding to a Change in One Element of a Given Matrix.} \\
        Annals of Mathematical Statistics 21, 1950. doi: \url{http://dx.doi.org/10.1214/aoms/1177729893}.
    \bibitem{linalgDavid}
        David C. Lay Chapter 3,
        \textit{Linear Algebra and Its Applications},
        Pearson, 2011, ISBN 978-0321835144.
    \bibitem{numstoch}
        Peter E. Kloeden and Eckhard Platen,
        \textit{Numerical Solution of Stochastic Differential Equations},
        Springer, 1992, ISBN 978-3-662-12616-5.
    \bibitem{statmech}
        J.P. Sethna,
        \textit{Statistical Mechanics: Entropy, Order Parameters and Complexity},
        \url{http://pages.physics.cornell.edu/sethna/StatMech/EntropyOrderParametersComplexity.pdf}.
\end{thebibliography}
